Optimizing a PoS Tag Set for Dependency Parsing

### 2016-05-04 submission.tex
Added some refs

\section{Introduction}
\label{sec:intro}
PoS tagging is an important preprocessing step for many NLP
tasks, such as parsing, named entity recognition and sentiment
analysis. Whereas much effort has gone into the development of
PoS-taggers, to the effect that this task is often considered
more or less a solved task, considerably less effort has been
devoted to the empirical evaluation of the PoS tag sets
themselves.
PoS tag sets are usually employed in a "one size fits
all" fashion, regardless of the requirements posed by the
devoted to empirical evaluation of the PoS tag sets
themselves.
PoS tag sets are usually employed in a ``one size fits
all'' fashion, regardless of the requirements posed by the
downstream NLP tasks for which PoS-tagging is performed.

It is well known that syntactic parsing often benefits from quite
fine-grained morphological distinctions since morphology interacts
with syntax though phenomena such as agreement and case
marking. However, in a realistic setting where the aim is to
automatically parse raw text, the generation of morphological
information will often require a separate step of morphological
analysis that can be quite costly.

In this paper, we optimize a Norwegian PoS tag set for the task of
data-driven dependency parsing.  We show that the introduction of
morphological distinctions not present in the original tag set, whilst
compromising tagger accuracy, actually leads to significantly improved
parsing accuracy.
data-driven dependency parsing. In a set of experiments, various
morphological distinctions are introduced and evaluated both
intrinsically, i.e. in terms of PoS-tagging accuracy and
extrinsically, in terms of parsing accuracy.  Our results show that
the introduction of morphological distinctions not present in the
original tag set, whilst compromising tagger accuracy, actually leads
to significantly improved parsing accuracy. This optimization allows
us to bypass the additional step of morphological analysis, framing
the whole preprocessing problem as a simple tagging task.

>>>>

\section{Introduction}
\label{sec:intro}
PoS tagging is an important preprocessing step for many NLP tasks,
such as parsing, named entity recognition and sentiment
analysis. Whereas much effort has gone into the development of
PoS-taggers, to the effect that this task is often considered more or
less a solved task, considerably less effort has been devoted to the
empirical evaluation of the PoS tag sets themselves. Error analysis of
PoS-taggers indicate that, whereas tagger improvement though means of
learning algorithm or feature engineering seems to have reached
something of a plateau, linguistic assessment of the distinctions made
by the PoS tag set may be a venue worth investigating further
\cite{Man:11}.  Clearly, the utility of a PoS tag set is tightly
coupled with the the downstream task for which it is performed. Even
so, PoS tag sets are usually employed in a ``one size fits all''
fashion, regardless of the requirements posed by the task which makes
use of this information.

It is well known that syntactic parsing often benefits from quite
fine-grained morphological distinctions \cite{Zha:Niv:11}. Morphology
interacts with syntax though phenomena such as agreement and case
marking and incorporating information on morphological properties of
words can therefore often improved parsing performance. However, in a
realistic setting where the aim is to automatically parse raw text,
the generation of morphological information will often require a
separate step of morphological analysis that can be quite costly.

In this paper, we optimize a Norwegian PoS tag set for the task of
data-driven dependency parsing.  We show that the introduction of
morphological distinctions not present in the original tag set, whilst
compromising tagger accuracy, actually leads to significantly improved
parsing accuracy.  In a set of experiments, various morphological
distinctions are introduced and evaluated both intrinsically, i.e. in
terms of PoS-tagging accuracy, and extrinsically, in terms of parsing
accuracy.  Our results show that the introduction of morphological
distinctions not present in the original tag set, whilst compromising
tagger accuracy, actually leads to significantly improved parsing
accuracy. This optimization allows us to bypass the additional step of
morphological analysis, framing the whole preprocessing problem as a
simple tagging task.
